---
title: "probiotic_metabolomics"
author: "HCW"
date: "11/21/2020"
output: html_document
---

```{r}
library(resample) 
library(tidyverse)
library(siggenes)
library(vegan)
library(nlme)
library(zoo)
library(Hmisc)
```

```{r}
save_figure_path <- "~/R/Project_Probiotic/CLEAN_CODE/plots/"

metab_data_raw <- read.csv("~/R/metabolomics_projects/Pilot_Probiotic_shwvt_library_update/probiotic_serum_metabolites_all_norm.csv") %>% 
  mutate(Participant=as.integer(str_sub(as.character(Participant),-2,-1))) %>% 
  dplyr::rename(CTRU_id=Participant)

prob_responders <- c(7041,7006,7027,7047,7002,7001,7050,7018,7009,7012,7015,7023,7005,7017)
project_probiotic_ppts <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/data/project_probiotic_ppts.csv") %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  mutate(Participant=as.integer(as.character(Participant)))
corrected_grouping <- read.csv("~/R/Project_Probiotic/CLEAN_CODE/data/41697_OLINK_serum_2018-11-16.csv",stringsAsFactors = FALSE) %>% 
  dplyr::rename(CTRU_id=CTRU..OpenSpecimen.ID,Participant=Project.Probiotic.Study.ID) %>% 
  select(CTRU_id,Participant) %>% 
  left_join(.,project_probiotic_ppts) %>% 
  unique()

metab_data <- metab_data_raw %>% 
  right_join(corrected_grouping,.) %>% 
  select(-CTRU_id,-X,-Sample_type,-Study,-AMILORIDEPEAK2_c18pos,-DILTIAZEM_c18pos,-X4.ACETAMIDOPHENOL_HILICpos,-LORATADINE_HILICpos) %>% 
  filter(Timepoint != 5) %>% 
  na.omit()
metab_data_spread_scale <- metab_data %>% 
  mutate(Participant=as.character(Participant),
         Timepoint=as.character(Timepoint)) %>% 
  mutate_if(is.numeric, scale)
#write.csv(metab_data_spread_scale,"~/R/metabolomics_projects/Pilot_Probiotic_shwvt_library_update/probiotic_serum_metabolites_clean_scale.csv")

metab_data_gather <- metab_data %>% 
  gather(key=metab,value=metab_value,-Participant,-treatment_type,-prob_response,-Timepoint)
```


```{r}
paired_df_ppt <- function(df,tp_1,tp_2){
  df <- df %>% 
    arrange(Participant)
  temp_1 <- df %>% 
    filter(Timepoint==tp_1) 
  temp_2 <- df %>% 
    filter(Timepoint==tp_2)
  ppt_list <- intersect(temp_1$Participant,temp_2$Participant)
  
  df_paired <- bind_rows(
    filter(temp_1, Participant %in% ppt_list),
    filter(temp_2,Participant %in% ppt_list)
  )
  return(df_paired)
}

var_function <- function(df_spread,num_quantile){
  vars_vector <- df_spread %>% 
    select(-Participant,-Timepoint,-treatment_type,-prob_response) %>% 
    as.matrix() %>% 
    colVars(na.rm = T)
  name_vector <- df_spread %>% 
    select(-Participant,-Timepoint,-treatment_type,-prob_response) %>% 
    colnames()
  filt_names <- name_vector[vars_vector>quantile(vars_vector)[num_quantile]] #top 75% varying proteins =2
  return(filt_names)
}
```

Paired siggenes end of intervention to baseline
-filtered to top 75% most varying metabolites (like the host proteins)
-no sig individual metabolites from baseline to end of intervention
```{r}
#probiotic
metab_data_probiotic <- metab_data %>% 
  filter(treatment_type=="Probiotic") 

metab_names <- var_function(metab_data_probiotic,2)
metab_data_probiotic_filt <- metab_data_gather %>% 
  filter(treatment_type=="Probiotic" & metab %in% metab_names) %>% 
  spread(key=metab,value=metab_value)

metab_data_probiotic_paired <- paired_df_ppt(metab_data_probiotic_filt,1,6)
set.seed(20)
siggenes_cl <- c((-1:-23), 1:23)
siggenes_data <- t(metab_data_probiotic_paired[-c(1:4)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)

#placebo
#probiotic
# metab_data_placebo <- metab_data %>% 
#   filter(treatment_type=="Placebo") 
# 
# metab_names <- var_function(metab_data_placebo,2)
# metab_data_placebo_filt <- metab_data_gather %>% 
#   filter(treatment_type=="Placebo" & metab %in% metab_names) %>% 
#   spread(key=metab,value=metab_value) 
# 
# metab_data_placebo_paired <- paired_df_ppt(metab_data_placebo_filt,1,6)
# set.seed(20)
# siggenes_cl <- c((-1:-12), 1:12)
# siggenes_data <- t(metab_data_placebo_paired[-c(1:4)])
# siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
# summary(siggenes_output)
# delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
# siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)


#probiotic R
metab_data_probiotic_R <- metab_data %>% 
  filter(prob_response=="Responder") 

metab_names <- var_function(metab_data_probiotic_R,2)
metab_data_probiotic_R_filt <- metab_data_gather %>% 
  filter(prob_response=="Responder" & metab %in% metab_names) %>% 
  spread(key=metab,value=metab_value)

metab_data_probiotic_R_paired <- paired_df_ppt(metab_data_probiotic_R_filt,1,6)
set.seed(20)
siggenes_cl <- c((-1:-13), 1:13)
siggenes_data <- t(metab_data_probiotic_R_paired[-c(1:4)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
```

Use scaled differences for analysis that uses all metabolite features at once 
```{r}
#find paired samples - particvipants with both baseline and maint samples
metab_data_scale_paired <- paired_df_ppt(metab_data_spread_scale,1,6)
metab_data_scale_paired_probiotic <- metab_data_scale_paired %>% 
  filter(treatment_type=="Probiotic")
metab_data_scale_paired_placebo <- metab_data_scale_paired %>% 
  filter(treatment_type=="Placebo")

#calcualte difference between baseline and maint
metab_data_scale_diff <- metab_data_scale_paired %>% 
  gather(key=metab,value=metab_scaled,-c(Participant:Timepoint)) %>% 
  spread(key=Timepoint,value=metab_scaled) %>% 
  mutate(metab_scale_diff=`6`-`1`) %>% 
  select(-`6`,-`1`) %>% 
  spread(key=metab,value=metab_scale_diff)


```

Beta diversity of scaled metabolite difference 
```{r}
set.seed(200)
#between probiotic and placebo
metab_data_scale_diff_raw <- metab_data_scale_diff %>% 
  select(-Participant,-prob_response,-treatment_type)
permanova <- adonis(metab_data_scale_diff_raw ~ treatment_type, data = metab_data_scale_diff, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["treatment_type", "Pr(>F)"]

#between probiotic R, NR, and placebo
permanova <- adonis(metab_data_scale_diff_raw ~ prob_response, data = metab_data_scale_diff, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["prob_response", "Pr(>F)"]

#between baseline and maint probiotic group 
metab_data_scale_paired_probiotic_raw <- metab_data_scale_paired_probiotic %>% 
  select(-Participant,-prob_response,-treatment_type,-Timepoint)
permanova <- adonis(metab_data_scale_paired_probiotic_raw ~ Timepoint, data = metab_data_probiotic_paired, method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint", "Pr(>F)"]

#between baseline and maint placebo group 
metab_data_scale_paired_placebo_raw <- metab_data_scale_paired_placebo %>% 
  select(-Participant,-prob_response,-treatment_type,-Timepoint)
permanova <- adonis(metab_data_scale_paired_placebo_raw ~ Timepoint, data = metab_data_scale_paired_placebo, method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint", "Pr(>F)"]

#between baseline and maint probiotic R group 
metab_data_scale_paired_probiotic_R_raw <- metab_data_scale_paired_probiotic %>% 
  filter(prob_response=="Responder") %>% 
  select(-Participant,-prob_response,-treatment_type,-Timepoint)
permanova <- adonis(metab_data_scale_paired_probiotic_R_raw ~ Timepoint, data =filter(metab_data_scale_paired_probiotic,prob_response=="Responder"), method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint", "Pr(>F)"]

#between baseline and maint probiotic NR group 
metab_data_scale_paired_probiotic_R_raw <- metab_data_scale_paired_probiotic %>% 
  filter(prob_response=="Non-responder") %>% 
  select(-Participant,-prob_response,-treatment_type,-Timepoint)
permanova <- adonis(metab_data_scale_paired_probiotic_R_raw ~ Timepoint, data =filter(metab_data_scale_paired_probiotic,prob_response=="Non-responder"), method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint", "Pr(>F)"]
```

Plot PCoA space from the adonis 
```{r}
dis <- vegdist(metab_data_scale_diff_raw,method="eu")
## Calculate multivariate dispersions
mod <- betadisper(dis, metab_data_scale_diff$prob_response)

betadisper_df <- data.frame(Participant=metab_data_scale_diff$Participant,
                            treatment_type=metab_data_scale_diff$treatment_type,
                            prob_response=metab_data_scale_diff$prob_response,
                            dist_cent=mod[["distances"]],
                            mod[["vectors"]]) 
# % var PCoA 1
mod[["eig"]][1]/sum(mod[["eig"]])
# % var PCoA 2
mod[["eig"]][2]/sum(mod[["eig"]])

ggplot(betadisper_df,aes(x=PCoA1,y=PCoA2,colour=prob_response))+
  geom_point(size=3)+
  stat_ellipse(aes(x=PCoA1,y=PCoA2,colour=prob_response),level = 0.50)+
  scale_color_manual(values=c("#b2182b","gray85","#2166ac"))+
  theme_classic()+
  xlab("Metabolite Profile PCoA1 (8.9%)")+
  ylab("Metabolite Profile PCoA2 (7.1%)")+
  theme(text = element_text(size=18))
#ggsave(paste(save_figure_path,"metabolite_pcoa.pdf"),width=10,height = 8)

t.test(filter(betadisper_df,treatment_type=="Probiotic")$dist_cent,filter(betadisper_df,treatment_type=="Placebo")$dist_cent)
t.test(filter(betadisper_df,prob_response=="Responder")$dist_cent,filter(betadisper_df,prob_response=="Non-responder")$dist_cent)

ggplot(betadisper_df,aes(x=prob_response,y=dist_cent,fill=prob_response))+
  geom_boxplot()+
  scale_fill_manual(values=c("#b2182b","gray85","#2166ac"))+
  theme_classic()
```

LOOCV predict probiotic vs. placebo
```{r}
loocv_rf_function <- function(df) {
  set.seed(20)
  folds <- groupKFold(df$Participant,k=length(unique(df$Participant)))
  control <- rfeControl(functions=rfFuncs, method="LOOCV",number = 100,index=folds)
  df_temp <- select(df,-Participant) %>% as.data.frame
  df_centered <- center_scale(df_temp[,2:dim(df_temp)[2]])
  rfe_obj <- rfe(df_centered,as.factor(df_temp[,1]), sizes=(2:ncol(df_temp)-1), rfeControl=control)
  return(rfe_obj)
}
center_scale <- function(x) {
    scale(x, scale = FALSE)
}
loocv_results_dfs <- function(loocv_object,data_type_char){
  accuracy_df <- data.frame(data_type=data_type_char,
                            max_acc=max(loocv_object[["results"]][["Accuracy"]]),
                            optsize=loocv_object[["optsize"]])
  optVariables_df <- data.frame(data_type=data_type_char,
                            rank_order=c(1:length(loocv_object[["optVariables"]])),
                            optVariables=loocv_object[["optVariables"]])
  return(list(accuracy_df,optVariables_df))
}
```

```{r}
df_in <- metab_data_scale_diff %>% 
  select(-prob_response) %>% 
  mutate(treatment_type=as.factor(treatment_type))

loocv_rf_int_metab_change <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_metab_change_pp <- loocv_rf_int_metab_change %>% 
  loocv_results_dfs(.,"metab_change_pp")

loocv_rf_results_metab_change_pp_sigg <- loocv_rf_results_metab_change_pp[[2]] %>% 
  filter(rank_order<5) 
metab_data_scale_diff_pp_sigg <- metab_data_scale_diff %>%
  gather(key=metab,value=metab_change,-Participant,-treatment_type,-prob_response) %>% 
  filter(metab %in% loocv_rf_results_metab_change_pp_sigg$optVariables)
ggplot(metab_data_scale_diff_pp_sigg,aes(x=treatment_type,y=metab_change,fill=treatment_type))+
  geom_boxplot()+
  theme_classic()+
  scale_fill_manual(values=c("gray85", "thistle4"))+
  facet_wrap(~metab,scales="free")

#probiotic R vs. NR
df_in <- metab_data_scale_diff %>% 
  filter(treatment_type=="Probiotic") %>% 
  select(-treatment_type) %>% 
  mutate(prob_response=as.factor(prob_response))

loocv_rf_int_metab_change_rnr <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_metab_change_rnr <- loocv_rf_int_metab_change_rnr %>% 
  loocv_results_dfs(.,"metab_change_rnr")

metab_data_scale_diff <- metab_data_scale_diff %>% 
  mutate(prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE))
ggplot(metab_data_scale_diff,aes(x=prob_response,y=HOMOVANILLIC.ACID_c18pos,fill=prob_response))+
  geom_boxplot()+
  theme_classic()+
  scale_fill_manual(values=c("gray","#b2182b","#2166ac"))+
  theme(text = element_text(size=18))+
  ylab("Scaled Difference in Serum Homovanillic Acid")+
  xlab("Treatment Arm")
# ggsave(paste(save_figure_path,"metabolite_homovanillic_acid.pdf"),width=10,height = 8)


t.test(filter(metab_data_scale_diff,prob_response=="Responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Non-responder")$HOMOVANILLIC.ACID_c18pos)
t.test(filter(metab_data_scale_diff,prob_response=="Responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Placebo")$HOMOVANILLIC.ACID_c18pos)
t.test(filter(metab_data_scale_diff,prob_response=="Non-responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Placebo")$HOMOVANILLIC.ACID_c18pos)
```


Elastic net predicting group based on metabolite change
```{r}
#probiotic vs. placebo 
df_EN_raw <- metab_data_scale_diff %>% 
  select(-Participant,-prob_response) %>% 
  mutate(treatment_type=as.factor(treatment_type))
set.seed(20)
trainIndex <- createDataPartition(df_EN_raw$treatment_type, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(treatment_type ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$treatment_type,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))

#probiotic R vs. NR 
df_EN_raw <- metab_data_scale_diff %>% 
  filter(treatment_type=="Probiotic") %>% 
  select(-Participant,-treatment_type) %>% 
  mutate(prob_response=as.factor(prob_response))
set.seed(20)
trainIndex <- createDataPartition(df_EN_raw$prob_response, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(prob_response ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$prob_response,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))
```

Elastic net predicting probiotic R vs. NR based on baseline metabolites
```{r}
#probiotic R vs. NR 
df_EN_raw <- metab_data_spread_scale %>% 
  filter(treatment_type=="Probiotic" & Timepoint==1) %>% 
  select(-Participant,-treatment_type,-Timepoint) %>% 
  mutate(prob_response=as.factor(prob_response))
set.seed(200)
trainIndex <- createDataPartition(df_EN_raw$prob_response, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(prob_response ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$prob_response,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))
```
adonis on baseline metabolite profile for probiotic R vs NR
```{r}
metab_data_spread_scale_base <- metab_data_spread_scale %>% 
  filter(treatment_type=="Probiotic" & Timepoint==1) 

metab_data_spread_scale_base_raw <- metab_data_spread_scale_base %>% 
  select(-Participant,-treatment_type,-Timepoint,-prob_response)

#between probiotic R, NR, and placebo
permanova <- adonis(metab_data_spread_scale_base_raw ~ prob_response, data = metab_data_spread_scale_base, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["prob_response", "Pr(>F)"]
```
Distance from centroid 
```{r}
metab_data_spread_scale_raw <- metab_data_spread_scale %>% 
  select(-Participant,-treatment_type,-Timepoint,-prob_response)

dis <- vegdist(metab_data_spread_scale_raw,method="eu")
## Calculate multivariate dispersions
mod <- betadisper(dis, paste(metab_data_spread_scale$prob_response,metab_data_spread_scale$Timepoint,sep="_"))

betadisper_df <- data.frame(Participant=metab_data_spread_scale$Participant,
                            treatment_type=metab_data_spread_scale$treatment_type,
                            prob_response=metab_data_spread_scale$prob_response,
                            Timepoint_label=metab_data_spread_scale$Timepoint,
                            response_time=paste(metab_data_spread_scale$prob_response,metab_data_spread_scale$Timepoint,sep="_"),
                            dist_cent=mod[["distances"]],
                            mod[["vectors"]]) 

ggplot(betadisper_df,aes(x=PCoA1,y=PCoA2,colour=response_time))+
  geom_point()+
  stat_ellipse(aes(x=PCoA1,y=PCoA2,colour=response_time),level = 0.50)+
  # scale_color_manual(values=c("#d95f02","gray85","#1b9e77"))+
  theme_classic()

t.test(filter(betadisper_df,response_time=="Responder_1")$dist_cent,filter(betadisper_df,response_time=="Responder_6")$dist_cent)
t.test(filter(betadisper_df,response_time=="Non-responder_1")$dist_cent,filter(betadisper_df,response_time=="Non-responder_6")$dist_cent)
t.test(filter(betadisper_df,response_time=="Non-responder_1")$dist_cent,filter(betadisper_df,response_time=="Responder_1")$dist_cent)

ggplot(filter(betadisper_df,response_time %in% c("Responder_Late_int","Non-responder_Late_int")),aes(x=response_time,y=dist_cent,fill=response_time))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d95f02","gray85","#1b9e77"))+
  theme_classic()

```

