---
title: "parse_raw_olink"
author: "HCW"
date: "12/10/2018"
output: html_document
---

Package Loading
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(FactoMineR)
library(ggfortify)
library(factoextra)
library(randomForest)
```

Import data
```{r}
save_figure_path <- "~/R/Project_Probiotic/CLEAN_CODE/plots/"

project_probiotic_ppts <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/data/project_probiotic_ppts.csv")
raw_olink <- read.csv("~/R/Project_Probiotic/CLEAN_CODE/data/20181072_Sonnenburg_NPX_LODP_edit.csv",stringsAsFactors = FALSE) %>% 
  filter(QC.Warning=="Pass")
temp <- tidyr::separate(raw_olink,col=Assay,into =c("runID","CTRU_ID","Temp_ID","tempa"),sep="-",remove=FALSE) %>% 
  dplyr::filter(.,tempa =="a"|is.na(tempa)) %>% 
  dplyr::select(.,-tempa) %>% 
  dplyr::filter(.,runID=="41697")%>% 
  tidyr::unite(.,"Assay",c("runID","CTRU_ID","Temp_ID"),sep="-")
olink_qc <- read.csv("~/R/Project_Probiotic/CLEAN_CODE/data/olink_QC.csv")
olink_keep_vector <- filter(olink_qc,Missing.Data.freq.<.25)$Assay
  
redcap_name_tp_number <- data.frame(Timepoint_str=c("Baseline -3", "Baseline", "Week 4","Week 6","Week 8","Week 10","Week 14"),
                                    Timepoint=c(2,3,5,6,7,8,10))
corrected_grouping <- read.csv("~/R/Project_Probiotic/CLEAN_CODE/data/41697_OLINK_serum_2018-11-16.csv",stringsAsFactors = FALSE) %>% 
  dplyr::rename(., Assay=Serum.Label,Timepoint_str=Timepoint,Timepoint=Visit..,Participant=Project.Probiotic.Study.ID,CTRU_ID=CTRU..OpenSpecimen.ID) %>% 
  select(-Timepoint) %>% 
  left_join(.,redcap_name_tp_number)

olink_data <- left_join(dplyr::select(corrected_grouping, Participant,Timepoint,Assay),temp,by="Assay") %>%
  dplyr::select(.,-Assay,-Plate.ID,-QC.Warning) %>% 
  .[complete.cases(.), ] %>% 
  gather(key=cytokine,value=cytokine_value,-Participant,-Timepoint) %>% 
  mutate(cytokine_value=as.numeric(cytokine_value),
         Participant=as.factor(Participant)) %>% 
  left_join(project_probiotic_ppts,.) %>% 
  mutate(Timepoint_label=ifelse(Timepoint %in% c(1,2,3),"Baseline",
                                ifelse(Timepoint %in% c(4,5,6),"Early_int",
                                       ifelse(Timepoint %in% c(8),"Late_int",
                                              "Washout")))) %>% 
  filter(Participant != 7050)
  # filter(cytokine %in% olink_keep_vector)

# prob_responders <- c(7041,7031,7001,7018,7002,7038,7050,7012,7009,7017,7046,7006,7019,7047,7023,7015,7026,7005)
# prob_responders <- c(7046,7006,7001,7050,7017,7018,7047,7009,7019,7026,7023,7015,7005)
# prob_responders <- c(7041,7031,7002,7006,7046,7001,7018,7009,7050,7023,7047,7017,7015,7026,7005)
prob_responders <- c(7041,7006,7027,7047,7002,7001,7050,7018,7009,7012,7015,7023,7005,7017)

write.csv(olink_data,"~/R/Project_Probiotic/CLEAN_CODE/data/olink_clean.csv")

```

Siggenes to see if any cytokines significantly changed from baseline to end of intervention
```{r}
paired_df_ppt <- function(df,tp_1,tp_2){
  df <- df %>% 
    arrange(Participant)
  temp_1 <- df %>% 
    filter(Timepoint==tp_1) 
  temp_2 <- df %>% 
    filter(Timepoint==tp_2)
  ppt_list <- intersect(temp_1$Participant,temp_2$Participant)
  
  df_paired <- bind_rows(
    filter(temp_1, Participant %in% ppt_list),
    filter(temp_2,Participant %in% ppt_list)
  )
  return(df_paired)
}

var_function <- function(df_spread,num_quantile){
  vars_vector <- df_spread %>% 
    select(-Participant,-Timepoint,-treatment_type,-Timepoint_label) %>% 
    as.matrix() %>% 
    colVars()
  name_vector <- df_spread %>% 
    select(-Participant,-Timepoint,-treatment_type,-Timepoint_label) %>% 
    colnames()
  filt_names <- name_vector[vars_vector>quantile(vars_vector)[num_quantile]] #top 75% varying proteins =2
  return(filt_names)
}
```

Paired siggenes end of intervention to baseline
-filtered to top 75% most varying cytokines
-no sig individual metabolites from baseline to end of intervention
```{r}
#Probiotic grop
olink_data_probiotic <- olink_data %>% 
  filter(treatment_type=="Probiotic") %>% 
  spread(key=cytokine,value=cytokine_value) 

olink_probiotic_names <- var_function(olink_data_probiotic,2)
olink_data_probiotic_filt <- olink_data %>% 
  filter(treatment_type=="Probiotic" & cytokine %in% olink_probiotic_names) %>% 
  spread(key=cytokine,value=cytokine_value)

ctyokine_probiotic_filt_paired <- paired_df_ppt(olink_data_probiotic_filt,3,8)
set.seed(20)
siggenes_cl <- c((-1:-22), 1:22)
siggenes_data <- t(ctyokine_probiotic_filt_paired[-c(1:4)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)


#probiotic responders group
olink_data_probiotic_resp <- olink_data %>% 
  filter(Participant %in% prob_responders) %>% 
  spread(key=cytokine,value=cytokine_value)

olink_probiotic_resp_names <- var_function(olink_data_probiotic_resp,2)
olink_data_probiotic_resp_filt <- olink_data %>% 
  filter(Participant %in% prob_responders & cytokine %in% olink_probiotic_resp_names) %>% 
  spread(key=cytokine,value=cytokine_value)

ctyokine_probiotic_resp_filt_paired <- paired_df_ppt(olink_data_probiotic_resp_filt,3,8)
set.seed(20)
siggenes_cl <- c((-1:-11), 1:11)
siggenes_data <- t(ctyokine_probiotic_resp_filt_paired[-c(1:4)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)
siggenes_df <- summary(siggenes_output,0.1)


#placebo grop
olink_data_placebo <- olink_data %>% 
  filter(treatment_type=="Placebo") %>% 
  spread(key=cytokine,value=cytokine_value)

olink_placebo_names <- var_function(olink_data_placebo,2)
olink_data_placebo_filt <- olink_data %>% 
  filter(treatment_type=="Placebo" & cytokine %in% olink_placebo_names) %>% 
  spread(key=cytokine,value=cytokine_value)

olink_data_placebo_filt_paired <- paired_df_ppt(olink_data_placebo_filt,3,8)
set.seed(20)
siggenes_cl <- c((-1:-12), 1:12)
siggenes_data <- t(olink_data_placebo_filt_paired[-c(1:4)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)
siggenes_df <- summary(siggenes_output,0.1)
```

Average the two baselin etimepoints and compare it to the end of intervention (late_int) levels
```{r}
paired_df_ppt <- function(df,tp_1,tp_2){
  df <- df %>% 
    arrange(Participant)
  temp_1 <- df %>% 
    filter(Timepoint_label==tp_1) 
  temp_2 <- df %>% 
    filter(Timepoint_label==tp_2)
  ppt_list <- intersect(temp_1$Participant,temp_2$Participant)
  
  df_paired <- bind_rows(
    filter(temp_1, Participant %in% ppt_list),
    filter(temp_2,Participant %in% ppt_list)
  )
  return(df_paired)
}
var_function <- function(df_spread,num_quantile){
  vars_vector <- df_spread %>% 
    select(-Participant,-treatment_type,-Timepoint_label) %>% 
    as.matrix() %>% 
    colVars()
  name_vector <- df_spread %>% 
    select(-Participant,-treatment_type,-Timepoint_label) %>% 
    colnames()
  filt_names <- name_vector[vars_vector>quantile(vars_vector)[num_quantile]] #top 75% varying proteins =2
  return(filt_names)
}

olink_data_tp_chunk <- olink_data %>% 
  group_by(Participant,treatment_type,Timepoint_label,cytokine) %>% 
  dplyr::summarize(cytokine_tp_avg=mean(cytokine_value,na.omit=TRUE)) %>% 
  ungroup
#write.csv(olink_data_tp_chunk,"~/R/Project_Probiotic/CLEAN_CODE/data/olink_data_tp_chunk.csv")

#Probiotic grop
olink_data_probiotic_tp_chunk <- olink_data_tp_chunk %>% 
  filter(treatment_type=="Probiotic") %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

olink_probiotic_names_tp_chunk <- var_function(olink_data_probiotic_tp_chunk,2)
olink_data_probiotic_filt_tp_chunk <- olink_data_tp_chunk %>% 
  filter(treatment_type=="Probiotic" & cytokine %in% olink_probiotic_names_tp_chunk) %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

ctyokine_probiotic_filt_paired_tp_chunk <- paired_df_ppt(olink_data_probiotic_filt_tp_chunk,"Baseline","Late_int")
set.seed(20)
siggenes_cl <- c((-1:-24), 1:24)
siggenes_data <- t(ctyokine_probiotic_filt_paired_tp_chunk[-c(1:3)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)

#Probiotic responders grop
olink_data_probiotic_tp_chunk <- olink_data_tp_chunk %>% 
  filter(Participant %in% prob_responders) %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

olink_probiotic_names_tp_chunk <- var_function(olink_data_probiotic_tp_chunk,2)
olink_data_probiotic_filt_tp_chunk <- olink_data_tp_chunk %>% 
  filter(Participant %in% prob_responders & cytokine %in% olink_probiotic_names_tp_chunk) %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

ctyokine_probiotic_filt_paired_tp_chunk <- paired_df_ppt(olink_data_probiotic_filt_tp_chunk,"Baseline","Late_int")
set.seed(20)
siggenes_cl <- c((-1:-13), 1:13)
siggenes_data <- t(ctyokine_probiotic_filt_paired_tp_chunk[-c(1:3)])
siggenes_output <- siggenes::sam(data=siggenes_data,cl=siggenes_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)
# siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)
# 
# olink_data_sig <- olink_data_tp_chunk %>% 
#   filter(cytokine=="CCL20" & Timepoint_label != "Early_int") %>% 
#   mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
#                               ifelse(Participant %in% prob_responders,"Responder","Non-responder")),
#          prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE))
# 
# ggplot(olink_data_sig,aes(x=Timepoint_label,y=cytokine_tp_avg,fill=prob_response))+
#   geom_boxplot()+
#   scale_fill_discrete(name = "prob_response", labels = c("Placebo","Non-responder","Responder"))+
#   scale_fill_manual(values=c("gray85", "thistle3","thistle4"))+
#   theme_classic()
```


```{r}
unpairedYSAM <- function(df1,df2){
  y = c(rep(0,dim(df1)[1]),rep(1,dim(df2)[1]))
  return(y)
}

unpairedXSAM <- function(df1,df2,startIDs,endIDs){
  x1 = t(df1[-c(startIDs:endIDs)])
  #print(head(x1))
  x2 = t(df2[-c(startIDs:endIDs)])
  #print(head(x2))
  x_all <- cbind(x1,x2) 
  #print(x_all)
  return(x_all)
}

var_function <- function(df_spread,num_quantile){
  vars_vector <- df_spread %>% 
    select(-Participant,-treatment_type) %>% 
    as.matrix() %>% 
    colVars()
  name_vector <- df_spread %>% 
    select(-Participant,-treatment_type) %>% 
    colnames()
  filt_names <- name_vector[vars_vector>quantile(vars_vector)[num_quantile]] #top 75% varying proteins =2
  return(filt_names)
}
```

Calculate percent difference from baseline to end of intervention
-none sig
```{r}
olink_data_tp_chunk_percen_diff <- olink_data_tp_chunk %>% 
  spread(key=Timepoint_label,value=cytokine_tp_avg) %>% 
  mutate(Baseline_lateInt_percen=(Late_int-Baseline)/Baseline) %>% 
  select(Participant,treatment_type,cytokine,Baseline_lateInt_percen) %>% 
  spread(key=cytokine,value=Baseline_lateInt_percen) %>% 
  na.omit()

olink_names <- var_function(olink_data_tp_chunk_percen_diff,2)
olink_data_tp_chunk_percen_diff_filt <- olink_data_tp_chunk_percen_diff %>% 
  gather(key=cytokine,value=cytokine_diff_percen,-Participant,-treatment_type) %>% 
  filter(cytokine %in% olink_names) %>% 
  spread(key=cytokine,value=cytokine_diff_percen) %>% 
  ungroup
olink_data_tp_chunk_percen_diff_probiotic <- olink_data_tp_chunk_percen_diff_filt %>% 
  filter(treatment_type=="Probiotic")
olink_data_tp_chunk_percen_diff_placebo <- olink_data_tp_chunk_percen_diff_filt %>% 
  filter(treatment_type=="Placebo")

set.seed(20)
siggenes_input_data <- unpairedXSAM(olink_data_tp_chunk_percen_diff_probiotic,olink_data_tp_chunk_percen_diff_placebo,1,2)
siggenes_input_cl <- unpairedYSAM(olink_data_tp_chunk_percen_diff_probiotic,olink_data_tp_chunk_percen_diff_placebo)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)
# siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>% #q-value didn't pass 
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)

olink_data_tp_chunk_percen_diff_probiotic_responders <- olink_data_tp_chunk_percen_diff_probiotic %>% 
  filter(Participant %in% prob_responders)
set.seed(20)
siggenes_input_data <- unpairedXSAM(olink_data_tp_chunk_percen_diff_probiotic_responders,olink_data_tp_chunk_percen_diff_placebo,1,2)
siggenes_input_cl <- unpairedYSAM(olink_data_tp_chunk_percen_diff_probiotic_responders,olink_data_tp_chunk_percen_diff_placebo)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.1)

#plot change in ccl20 from the paired analysis in chunk above
olink_data_tp_chunk_percen_diff <- olink_data_tp_chunk_percen_diff %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder")),
         prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE))

# ggplot(olink_data_tp_chunk_percen_diff,aes(x=prob_response,y=CCL20,fill=prob_response))+
#   geom_boxplot()+
#   scale_fill_discrete(name = "prob_response", labels = c("Placebo","Non-responder","Responder"))+
#   scale_fill_manual(values=c("gray85", "thistle3","thistle4"))+
#   theme_classic()+
#   theme(text = element_text(size=18))+
#   ylab("Percent change from Baseline to End of Intervention")+
#   xlab("Probiotic Response Group")
# ggsave(paste(save_figure_path,"ccl20_percen_change_boxplot.pdf"),width=10,height = 8)

```

Calculate levels at end of intervention between probiotic and placebo groups
```{r}
olink_data_tp_chunk_late_int <- olink_data_tp_chunk %>% 
  filter(Timepoint_label=="Late_int") %>% 
  spread(key=cytokine,value=cytokine_tp_avg) %>% 
  na.omit() %>% 
  select(-Timepoint_label)

olink_names <- var_function(olink_data_tp_chunk_late_int,2)
olink_data_tp_chunk_late_int_filt <- olink_data_tp_chunk_late_int %>% 
  gather(key=cytokine,value=cytokine_tp_avg,-Participant,-treatment_type) %>% 
  filter(cytokine %in% olink_names) %>% 
  spread(key=cytokine,value=cytokine_tp_avg) %>% 
  ungroup
olink_data_tp_chunk_late_int_filt_prob <- olink_data_tp_chunk_percen_diff_filt %>% 
  filter(treatment_type=="Probiotic")
olink_data_tp_chunk_late_int_filt_plac <- olink_data_tp_chunk_percen_diff_filt %>% 
  filter(treatment_type=="Placebo")

set.seed(20)
siggenes_input_data <- unpairedXSAM(olink_data_tp_chunk_late_int_filt_prob,olink_data_tp_chunk_late_int_filt_plac,1,2)
siggenes_input_cl <- unpairedYSAM(olink_data_tp_chunk_late_int_filt_prob,olink_data_tp_chunk_late_int_filt_plac)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
```


Predict probiotic vs. placebo based on cytokine change
```{r}
olink_data_tp_chunk_percen_diff <- olink_data_tp_chunk_percen_diff %>% 
  mutate(prob_response=ifelse(Participant %in% prob_responders,"Responder","Non-responder"))

#probiotic vs. placebo 
df_EN_raw <- olink_data_tp_chunk_percen_diff %>% 
  select(-Participant,-prob_response) %>% 
  mutate(treatment_type=as.factor(treatment_type))
set.seed(2)
trainIndex <- createDataPartition(df_EN_raw$treatment_type, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(treatment_type ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$treatment_type,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))

#probiotic R vs. NR
df_EN_raw <- olink_data_tp_chunk_percen_diff %>%
  filter(treatment_type=="Probiotic") %>%
  select(-Participant,-treatment_type) %>% 
  mutate(prob_response=droplevels(as.factor(prob_response)))
set.seed(2)
trainIndex <- createDataPartition(df_EN_raw$prob_response, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(prob_response ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$prob_response,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))
```


Predict probiotic resposne based on baseline cytokine data
```{r}
olink_data_tp_chunk_prob_base <- olink_data_tp_chunk %>% 
  filter(Timepoint_label=="Baseline" & treatment_type=="Probiotic") %>% 
  mutate(prob_response=ifelse(Participant %in% prob_responders,"Responder","Non-responder")) %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

# df_in <- bind_rows(food_data_tp_chunk_resp_base,food_data_tp_chunk_NONresp_base)
# df_in <- bind_rows(ndsr_nut_resp_int,ndsr_nut_NONresp_int)

df_EN_raw <- olink_data_tp_chunk_prob_base %>% 
  select(-Participant,-treatment_type,-Timepoint_label) %>% 
  mutate(prob_response=as.factor(prob_response))
set.seed(2)
trainIndex <- createDataPartition(df_EN_raw$prob_response, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_EN_train <- df_EN_raw[ trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
df_EN_test  <- df_EN_raw[-trainIndex,] #%>% scale(center = TRUE, scale = TRUE)
set.seed(20)
enetFit <- train(prob_response ~ .,
                 data = df_EN_train,
                 method = "glmnet",
                 trControl=trainControl("LOOCV"),
                 family = "binomial")

coefs_matrix <- as.matrix(coef(enetFit$finalModel, enetFit$bestTune$lambda))
coefs_EN <- data.frame(metab=rownames(coefs_matrix),coef=coefs_matrix) %>%
  dplyr::rename(coef=X1) %>%
  filter(coef!=0 & metab!="(Intercept)") %>% 
  mutate(sign=coef>0)

pred_df_all <- data.frame(actual=df_EN_raw$prob_response,
                          predicted=predict(enetFit, newdata = df_EN_raw),
                          index_num=c(1:dim(df_EN_raw)[1]))
pred_df_all$tt <- ifelse(pred_df_all$index_num %in% as.vector(trainIndex),"train","test")
table(select(filter(pred_df_all,tt=="test"),actual,predicted))
```
Plot significant coefs from EN
```{r}
olink_data_tp_chunk_prob_base_gather_sigg <- olink_data_tp_chunk_prob_base %>% 
  gather(key=cytokine,value=base_value,-Participant,-treatment_type,-Timepoint_label,-prob_response) %>% 
  filter(cytokine %in% coefs_EN$metab)

ggplot(olink_data_tp_chunk_prob_base_gather_sigg,aes(x=prob_response,y=base_value,fill=prob_response))+
  geom_boxplot()+
  facet_wrap(~cytokine,scales="free")+
  theme_classic()+
  scale_fill_manual(values=c("#b2182b","#2166ac"))

```

```{r}
set.seed(200)
#between probiotic and placebo
olink_data_scale_diff_raw <- olink_data_tp_chunk_percen_diff %>% 
  select(-Participant,-prob_response,-treatment_type)
permanova <- adonis(olink_data_scale_diff_raw ~ treatment_type, data = olink_data_tp_chunk_percen_diff, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["treatment_type", "Pr(>F)"]

#between probiotic R, NR, and placebo
permanova <- adonis(olink_data_scale_diff_raw ~ prob_response, data = olink_data_tp_chunk_percen_diff, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["prob_response", "Pr(>F)"]

#between baseline and maint probiotic group 
olink_data_scale_paired_probiotic_raw <- ctyokine_probiotic_filt_paired_tp_chunk %>% 
  select(-Participant,-treatment_type,-Timepoint_label)
permanova <- adonis(olink_data_scale_paired_probiotic_raw ~ Timepoint_label, data = ctyokine_probiotic_filt_paired_tp_chunk, method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint_label", "Pr(>F)"]

#between baseline and maint probiotic R group 
olink_data_scale_paired_probiotic_R <- ctyokine_probiotic_filt_paired_tp_chunk %>% 
  mutate(prob_response=ifelse(Participant %in% prob_responders,"Responder","Non-responder")) %>% 
  filter(prob_response=="Responder") 
olink_data_scale_paired_probiotic_R_raw <- olink_data_scale_paired_probiotic_R %>% 
  select(-Participant,-prob_response,-treatment_type,-Timepoint_label)
permanova <- adonis(olink_data_scale_paired_probiotic_R_raw ~ Timepoint_label, data =olink_data_scale_paired_probiotic_R, method='euclidean',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["Timepoint_label", "Pr(>F)"]

#betrween probiotic and placebo, end of intervention only
olink_data_scale_late_int_raw <- olink_data_tp_chunk_late_int %>% 
  select(-Participant,-treatment_type)
permanova <- adonis(olink_data_scale_late_int_raw ~ treatment_type, data = olink_data_tp_chunk_late_int, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["treatment_type", "Pr(>F)"]

#between R vs. NR, end of intervention only
olink_data_tp_chunk_late_int_RNR <- olink_data_tp_chunk_late_int %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder")),
         prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE))

permanova <- adonis(olink_data_scale_late_int_raw ~ prob_response, data = olink_data_tp_chunk_late_int_RNR, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["prob_response", "Pr(>F)"]

#R vs. all at end of intervention only
olink_data_tp_chunk_late_int_RvsAll <- olink_data_tp_chunk_late_int_RNR %>%
  mutate(R_all=ifelse(prob_response=="Responder","R","all"))
permanova <- adonis(olink_data_scale_late_int_raw ~ R_all, data = olink_data_tp_chunk_late_int_RvsAll, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["R_all", "Pr(>F)"]
```

Plot PCoA space from the adonis 
```{r}
olink_data_tp_chunk_percen_diff <- olink_data_tp_chunk_percen_diff %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",ifelse(Participant %in% prob_responders,"Responder","Non-responder")))
olink_data_tp_chunk_percen_diff_raw <- olink_data_tp_chunk_percen_diff %>% 
  select(-Participant,-treatment_type,-prob_response)

dis <- vegdist(olink_data_scale_diff_raw,method="eu")
## Calculate multivariate dispersions
mod <- betadisper(dis, olink_data_tp_chunk_percen_diff$prob_response)

#get % variance
mod[["eig"]][1]/sum(mod[["eig"]])
# % var PCoA 2
mod[["eig"]][2]/sum(mod[["eig"]])

betadisper_df <- data.frame(Participant=olink_data_tp_chunk_percen_diff$Participant,
                            treatment_type=olink_data_tp_chunk_percen_diff$treatment_type,
                            prob_response=olink_data_tp_chunk_percen_diff$prob_response,
                            dist_cent=mod[["distances"]],
                            mod[["vectors"]]) 
ggplot(betadisper_df,aes(x=PCoA1,y=PCoA2,colour=prob_response))+
  geom_point(size=3)+
  stat_ellipse(aes(x=PCoA1,y=PCoA2,colour=prob_response),level = 0.50)+
  scale_color_manual(values=c("#b2182b","gray85","#2166ac"))+
  theme_classic()+
  xlab("Inflammatory Marker Profile PCoA1")+
  ylab("Inflammatory Marker Profile PCoA2")+
  theme(text = element_text(size=18))
ggsave(paste(save_figure_path,"cytokine_pcoa.pdf"),width=10,height = 8)

t.test(filter(betadisper_df,treatment_type=="Probiotic")$dist_cent,filter(betadisper_df,treatment_type=="Placebo")$dist_cent)
t.test(filter(betadisper_df,prob_response=="Responder")$dist_cent,filter(betadisper_df,prob_response=="Non-responder")$dist_cent)

ggplot(betadisper_df,aes(x=prob_response,y=dist_cent,fill=prob_response))+
  geom_boxplot()+
  scale_fill_manual(values=c("#b2182b","gray85","#2166ac"))+
  theme_classic()
```

LOOCV predict probiotic vs. placebo
```{r}
loocv_rf_function <- function(df) {
  set.seed(20)
  folds <- groupKFold(df$Participant,k=length(unique(df$Participant)))
  control <- rfeControl(functions=rfFuncs, method="LOOCV",number = 100,index=folds)
  df_temp <- select(df,-Participant) %>% as.data.frame
  df_centered <- center_scale(df_temp[,2:dim(df_temp)[2]])
  rfe_obj <- rfe(df_centered,as.factor(df_temp[,1]), sizes=(2:ncol(df_temp)-1), rfeControl=control)
  return(rfe_obj)
}
center_scale <- function(x) {
    scale(x, scale = FALSE)
}
loocv_results_dfs <- function(loocv_object,data_type_char){
  accuracy_df <- data.frame(data_type=data_type_char,
                            max_acc=max(loocv_object[["results"]][["Accuracy"]]),
                            optsize=loocv_object[["optsize"]])
  optVariables_df <- data.frame(data_type=data_type_char,
                            rank_order=c(1:length(loocv_object[["optVariables"]])),
                            optVariables=loocv_object[["optVariables"]])
  return(list(accuracy_df,optVariables_df))
}
```

```{r}
df_in <- olink_data_tp_chunk_percen_diff %>% 
  select(-prob_response) %>% 
  mutate(treatment_type=as.factor(treatment_type))

loocv_rf_int_olink_percen <- df_in %>% 
  loocv_rf_function()
loocv_rf_int_olink_percen_pp <- loocv_rf_int_olink_percen %>% 
  loocv_results_dfs(.,"olink_pp")

# loocv_rf_results_metab_change_pp_sigg <- loocv_rf_results_metab_change_pp[[2]] %>% 
#   filter(rank_order<5) 
# metab_data_scale_diff_pp_sigg <- metab_data_scale_diff %>%
#   gather(key=metab,value=metab_change,-Participant,-treatment_type,-prob_response) %>% 
#   filter(metab %in% loocv_rf_results_metab_change_pp_sigg$optVariables)
# ggplot(metab_data_scale_diff_pp_sigg,aes(x=treatment_type,y=metab_change,fill=treatment_type))+
#   geom_boxplot()+
#   theme_classic()+
#   scale_fill_manual(values=c("gray85", "thistle4"))+
#   facet_wrap(~metab,scales="free")

#probiotic R vs. NR
df_in <- olink_data_tp_chunk_percen_diff %>% 
  filter(treatment_type=="Probiotic") %>% 
  select(Participant,prob_response,ADA:X4E.BP1) %>% 
  mutate(prob_response=as.factor(prob_response))

loocv_rf_int_olink_percen_rnr <- df_in %>% 
  loocv_rf_function()
loocv_rf_int_olink_percen_rnr <- loocv_rf_int_olink_percen %>% 
  loocv_results_dfs(.,"olink_rnr")
# metab_data_scale_diff <- metab_data_scale_diff %>% 
#   mutate(prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE))
# ggplot(metab_data_scale_diff,aes(x=prob_response,y=HOMOVANILLIC.ACID_c18pos,fill=prob_response))+
#   geom_boxplot()+
#   theme_classic()+
#   scale_fill_manual(values=c("gray","#b2182b","#2166ac"))
# 
# t.test(filter(metab_data_scale_diff,prob_response=="Responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Non-responder")$HOMOVANILLIC.ACID_c18pos)
# t.test(filter(metab_data_scale_diff,prob_response=="Responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Placebo")$HOMOVANILLIC.ACID_c18pos)
# t.test(filter(metab_data_scale_diff,prob_response=="Non-responder")$HOMOVANILLIC.ACID_c18pos,filter(metab_data_scale_diff,prob_response=="Placebo")$HOMOVANILLIC.ACID_c18pos)
```
```{r}
olink_data_tp_chunk_spread <- olink_data_tp_chunk %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  spread(key=cytokine,value=cytokine_tp_avg)

olink_data_tp_chunk_spread_raw <- olink_data_tp_chunk_spread %>% 
  select(-Participant,-treatment_type,-Timepoint_label,-prob_response)

dis <- vegdist(olink_data_tp_chunk_spread_raw,method="eu")
## Calculate multivariate dispersions
mod <- betadisper(dis, paste(olink_data_tp_chunk_spread$prob_response,olink_data_tp_chunk_spread$Timepoint_label,sep="_"))

betadisper_df <- data.frame(Participant=olink_data_tp_chunk_spread$Participant,
                            treatment_type=olink_data_tp_chunk_spread$treatment_type,
                            prob_response=olink_data_tp_chunk_spread$prob_response,
                            Timepoint_label=olink_data_tp_chunk_spread$Timepoint_label,
                            response_time=paste(olink_data_tp_chunk_spread$prob_response,olink_data_tp_chunk_spread$Timepoint_label,sep="_"),
                            dist_cent=mod[["distances"]],
                            mod[["vectors"]]) 

ggplot(betadisper_df,aes(x=PCoA1,y=PCoA2,colour=response_time))+
  geom_point()+
  stat_ellipse(aes(x=PCoA1,y=PCoA2,colour=response_time),level = 0.50)+
  # scale_color_manual(values=c("#d95f02","gray85","#1b9e77"))+
  theme_classic()

t.test(filter(betadisper_df,response_time=="Responder_Late_int")$dist_cent,filter(betadisper_df,response_time=="Non-responder_Late_int")$dist_cent)
t.test(filter(betadisper_df,response_time=="Placebo_Baseline")$dist_cent,filter(betadisper_df,response_time=="Responder_Baseline")$dist_cent)

ggplot(filter(betadisper_df,response_time %in% c("Responder_Late_int","Non-responder_Late_int")),aes(x=response_time,y=dist_cent,fill=response_time))+
  geom_boxplot()+
  # scale_fill_manual(values=c("#d95f02","gray85","#1b9e77"))+
  theme_classic()

#baseline levels sig different between groups 
olink_data_tp_chunk_spread_base <- olink_data_tp_chunk_spread %>% 
  filter(Timepoint_label=="Baseline")
olink_data_tp_chunk_spread_base_raw <- olink_data_tp_chunk_spread_base %>% 
  select(-Participant,-treatment_type,-Timepoint_label,-prob_response)

permanova <- adonis(olink_data_tp_chunk_spread_base_raw ~ prob_response, data = olink_data_tp_chunk_spread_base, method='eu',na.rm = TRUE)
as.data.frame(permanova$aov.tab)["prob_response", "Pr(>F)"]
```

