---
title: "probiotic_food_groups"
author: "HCW"
date: "7/22/2019"
output: html_document
---

```{r}
library(randomForest)
library(tidyverse)
library(resample)
```

Import food data
```{r}
save_figure_path <- "~/R/Project_Probiotic/CLEAN_CODE/plots/"

project_probiotic_ppts <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/data/project_probiotic_ppts.csv")
redcap_name_tp_number <- data.frame(redcap_event_name=c("baseline_4_arm_1","baseline_2_arm_1","baseline_arm_1","week_2_arm_1","week_4_arm_1","week_6_arm_1","week_8_arm_1","week_10_arm_1","week_12_arm_1","week_14_arm_1"),
                                    Timepoint=c(1,2,3,4,5,6,7,8,9,10))
met_syn_all <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/data/met_syn_all.rds")

# prob_responders <- c(7041,7031,7001,7018,7002,7038,7050,7012,7009,7017,7046,7006,7019,7047,7023,7015,7026,7005)
# prob_responders <- c(7046,7006,7001,7050,7017,7018,7047,7009,7019,7026,7023,7015,7005)
# prob_responders <- c(7041,7031,7002,7006,7046,7001,7018,7009,7050,7023,7047,7017,7015,7026,7005)
prob_responders <- c(7041,7006,7027,7047,7002,7001,7050,7018,7009,7012,7015,7023,7005,7017) #with tp 1,2,3,7,8

# 
# prob_responders <- c(7005,7017,7023,7012,7015,7001,7050,7018,7002,7041,7006,7047,7027)

food_data_raw <- read.csv("~/R/Project_Probiotic/food_data_probiotic_study.csv") %>% 
  dplyr::rename(Participant=record_id) %>% 
  mutate(Participant=as.factor(Participant)) %>% 
  right_join(redcap_name_tp_number,.) %>% 
  inner_join(project_probiotic_ppts,.) %>% 
  filter(Energy..kcal.>=500)

food_data <- food_data_raw %>%
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  gather(key=food,value=food_value,-Participant,-treatment_type,-redcap_event_name,-Timepoint,-prob_response) %>% 
  select(-redcap_event_name) 

food_data_spread <- food_data %>% 
  spread(key=food,value=food_value)

#NDSR food data
ndsr_nutrients <- read.csv("~/R/Project_Probiotic/NDSR_food_data/NDSR_report_4_clean.csv") %>% 
  dplyr::rename(Participant=record_id) %>% 
  mutate(Timepoint_label=ifelse(Site.ID=="W0","Baseline","Intervention"),
         prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  select(-X,-hw360_id,-treatment_grp,-Site.ID,-c(Header.Data.Field.1.Descriptor:Trailer.Data.Field.3.Response),
         -c(Column.intentionally.left.blank:Column.intentionally.left.blank.3),-Trailer.Notes,-c(User.Nutrient.1..mg.:User.Nutrient.9..mg.),
         -c(User.Nutrient.10..mg.:User.Nutrient.19..mg.),-Data.Generated.in.NCC.Database.Version,-Data.Generated.in.Software.Version,-Header.Notes) %>% 
  gather(key=ndsr_nut,value=ndsr_value,-Participant,-treatment_type,-Timepoint_label,-prob_response) %>% 
  group_by(Participant,treatment_type,prob_response,Timepoint_label,ndsr_nut) %>% 
  dplyr::summarize(ndsr_value=mean(ndsr_value,na.rm=TRUE))
ndsr_nutrients_spread <- ndsr_nutrients %>% 
  spread(key=ndsr_nut,value=ndsr_value) %>% 
  ungroup


ndsr_food <- read.csv("~/R/Project_Probiotic/NDSR_food_data/NDSR_report_9_clean.csv") %>% 
  dplyr::rename(Participant=record_id) %>% 
  mutate(Timepoint_label=ifelse(Site.ID=="W0","Baseline","Intervention"),
         prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  select(-X,-hw360_id,-treatment_grp,-Site.ID,-Date.of.Intake,-Project.Abbreviation) %>% 
  gather(key=ndsr_food,value=ndsr_value,-Participant,-treatment_type,-Timepoint_label,-prob_response) %>% 
  group_by(Participant,treatment_type,prob_response,Timepoint_label,ndsr_food) %>% 
  dplyr::summarize(ndsr_value=mean(ndsr_value))
ndsr_food_spread <- ndsr_food %>% 
  spread(key=ndsr_food,value=ndsr_value) %>% 
  ungroup

ndsr_food_key <- read.csv("~/R/Project_Probiotic/NDSR_food_data/NDSR_report_9_food_key.csv")

ndsr_2_intake <- read.csv("~/R/Project_Probiotic/NDSR_food_data/NDSR_report_2_clean.csv") 

ndsr_2 <- ndsr_2_intake %>% 
  dplyr::rename(Participant=record_id) %>% 
  select(-X,-hw360_id,-treatment_grp,-Date.of.Intake,-Project.Abbreviation,-Food.File.ID,-Foods.Report.Item.Number,-c(Food.ID:Preparation.Food.Description)) %>%
    mutate(Timepoint_label=ifelse(Site.ID=="W0","Baseline","Intervention"),
         prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  select(-Site.ID) %>% 
  gather(key=food,value=food_value,-Participant,-Timepoint_label,-prob_response,-treatment_type,-Food.Name) %>% 
  mutate(food_value=as.numeric(food_value)) %>% 
  na.omit()

#take average of intake between W0 and W10 for each ppt
ndsr_2_tp_avg <- ndsr_2 %>% 
  group_by(Participant,treatment_type,prob_response,Food.Name,food) %>% 
  dplyr::summarize(food_value=mean(food_value,na.rm=TRUE))
ndsr_2_tp_med <- ndsr_2 %>% 
  group_by(Participant,treatment_type,prob_response,Food.Name,food) %>% 
  dplyr::summarize(food_value=median(food_value,na.rm=TRUE))
```

```{r}
weight_med <- met_syn_all %>% 
  select(Participant,Timepoint,treatment_type,WEIGHT) %>% 
  # mutate(tp_chunk=ifelse(Timepoint %in% c(1:3),"Baseline",ifelse(Timepoint %in% c(4:8),"Study","Washout"))) %>% 
  mutate(tp_chunk=ifelse(Timepoint %in% c(1:8),"Study","Washout")) %>% 
  group_by(Participant,treatment_type,tp_chunk) %>% 
  dplyr::summarize(weight_med=median(WEIGHT,na.rm=TRUE))

food_data_norm <- food_data %>% 
  # mutate(tp_chunk=ifelse(Timepoint %in% c(1:3),"Baseline",ifelse(Timepoint %in% c(4:8),"Study","Washout"))) %>% 
  mutate(tp_chunk=ifelse(Timepoint %in% c(1:8),"Study","Washout")) %>% 
  right_join(weight_med,.) %>% 
  mutate(food_value_norm=food_value/weight_med)

food_data_norm_tp_chunk <- food_data_norm %>% 
  group_by(Participant,treatment_type,prob_response,tp_chunk,food) %>% 
  dplyr::summarize(food_value_norm_med=median(food_value_norm,na.rm=TRUE))
  
t.test(filter(food_data_norm_tp_chunk,prob_response=="Responder" & tp_chunk=="Study" & food=="Energy..kcal.")$food_value_norm_med,
       filter(food_data_norm_tp_chunk,prob_response=="Non-responder" & tp_chunk=="Study" & food=="Energy..kcal.")$food_value_norm_med)
```

Stable weight and nutrition intake over course of study using LME across time
```{r}
LME_one_by_one_function <- function(df,names_vector,y_value_col,group_id_col){
  # df=temp_LME
  # names_vector=unique(met_syn_food_all_tps$food)
  # y_value_col=i
  # group_id_col="treatment_type"
  pValueList <- c()
  pValue_grpList <- c()
  xList <- c()
  correValueList <- c()
  stdEList <-c()
  intList <- c()
  for (i in names_vector){
    print(i)
    newdf <- data.frame(out = scale(df[,y_value_col]), 
                        xColName =scale(df[, i]), 
                        id = df[,"Participant"],
                        group_id=as.factor(df[,group_id_col]))
    #print(newdf)
    colnames(newdf)[2] <- "xColName"
    lmeData <- lme(xColName ~ out+group_id, data = newdf,random = ~1|id,na.action = na.omit,control=(msMaxIter=100))
    pVal = summary(lmeData)$tTable[2,5]
    group_pval = summary(lmeData)$tTable[3,5]
    corre = summary(lmeData)$tTable[2,1]
    stdE = summary(lmeData)$tTable[2,2]
    intercept = summary(lmeData)$tTable[1,1]
    xList <- c(xList,i)
    pValueList <- c(pValueList,pVal)
    pValue_grpList <- c(pValue_grpList,group_pval)
    correValueList <- c(correValueList,corre)
    stdEList <- c(stdEList,stdE)
    intList <- c(intList,intercept)
  }
  LME_pTable <- data.frame(xColName = xList,
                           correlation = correValueList,
                           pValue = pValueList,
                           pValueAdj = p.adjust(pValueList,method="fdr"),
                           pValue_grp=pValue_grpList,
                           pValue_grpAdj=p.adjust(pValue_grpList,method="fdr"),
                           intercept=intList,stdE=stdEList)
  return(LME_pTable)
}

food_list <- food_data %>% filter(!food %in% c("Formononetin..mg."))
food_time_lme_obj <- LME_one_by_one_function(food_data_spread %>% filter(Timepoint %in% c(3:8)),unique(food_list$food),"Timepoint","treatment_type") %>% 
  arrange(pValueAdj) %>% 
  dplyr::rename(Nutrition_intake=xColName)
write.csv(food_time_lme_obj,"~/R/Project_Probiotic/CLEAN_CODE/data/food_time_lme_obj_supp_table.csv")

weight_lme <- lme(WEIGHT ~ Timepoint+treatment_type, data = met_syn_all %>% filter(Timepoint %in% c(3:8)),random = ~1|Participant,na.action = na.omit,control=(msMaxIter=100))
summary(weight_lme)

ggplot(food_data_spread %>% mutate(Timepoint=as.factor(Timepoint)),aes(x=Timepoint,y=Energy..kcal.,fill=treatment_type))+
  geom_boxplot()+
  theme_classic()+
  xlab("Time (weeks)")+
  ylab("Energy (kcal)")+
  scale_fill_manual(values=c("gray85", "thistle4"))+
  scale_x_discrete(breaks=c(1,2,3,4,5,6,7,8,9,10),labels=c("-4","-2","0","2", "4","6", "8","10","12","14"),name="Time (weeks)")+
  theme(text = element_text(size=16))
#ggsave(paste(save_figure_path,"calories_time_study.pdf"),width=7 ,height = 5)

ggplot(met_syn_all %>% mutate(Timepoint=as.factor(Timepoint)),aes(x=Timepoint,y=WEIGHT,fill=treatment_type))+
  geom_boxplot()+
  theme_classic()+
  xlab("Time (weeks)")+
  ylab("Weight (kg)")+
  scale_fill_manual(values=c("gray85", "thistle4"))+
  scale_x_discrete(breaks=c(1,2,3,4,5,6,7,8,9,10),labels=c("-4","-2","0","2", "4","6", "8","10","12","14"),name="Time (weeks)")+
  theme(text = element_text(size=16))
#ggsave(paste(save_figure_path,"weight_time_study.pdf"),width=7 ,height = 5)

```


Baseline metab parameters predictors for R vs. NR
```{r}
metab_data_scale <- read.csv("~/R/metabolomics_projects/Pilot_Probiotic_shwvt_library_update/probiotic_serum_metabolites_clean_scale.csv") %>% 
  select(-X) %>% 
  mutate(Participant=as.character(Participant))
olink_data_tp_chunk <- read.csv("~/R/Project_Probiotic/CLEAN_CODE/data/olink_data_tp_chunk.csv") %>% 
  select(-X) %>% 
  spread(key=cytokine,value=cytokine_tp_avg) %>% 
  mutate(Participant=as.character(Participant),
         prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder")))

diversity_all <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/rds_obj/diversity_all.rds")
diversity_all_tp_chunk <- diversity_all %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  gather(key=diversity_meas,value=diversity_value,Observed:PD) %>% 
  mutate(Timepoint_label=ifelse(Timepoint %in% c(1:3),"Baseline",ifelse(Timepoint %in% c(4:8),"Study","Washout"))) %>%
  group_by(Participant,treatment_type,prob_response,Timepoint_label,diversity_meas) %>% 
  dplyr::summarize(value_tp_chunk=mean(diversity_value,na.rm=TRUE)) %>% 
  spread(key=diversity_meas,value=value_tp_chunk)

food_data_study <- food_data_norm %>% 
  # mutate(Timepoint_label=ifelse(Timepoint %in% c(1:3),"Baseline",ifelse(Timepoint %in% c(4:8),"Study","Washout"))) %>%
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder"))) %>% 
  mutate(Timepoint_label=ifelse(Timepoint %in% c(1:8),"Study","Washout")) %>% 
  group_by(Participant,prob_response,treatment_type,Timepoint_label,food) %>% 
  dplyr::summarize(value_tp_chunk=mean(food_value_norm,na.rm=TRUE)) %>% 
  spread(key=food,value=value_tp_chunk) %>% 
  filter(Timepoint_label=="Study") %>%
  mutate(Participant=as.character(Participant))
#merge all data frames at baselinefor prediction
metab_data_scale_base <- metab_data_scale %>% 
  filter(Timepoint==1) %>% 
  select(-Timepoint)
olink_data_base <- olink_data_tp_chunk %>% 
  filter(Timepoint_label=="Baseline")
diversity_all_base <- diversity_all_tp_chunk %>% 
  filter(Timepoint_label=="Baseline") %>% 
  mutate(prob_response=ifelse(treatment_type=="Placebo","Placebo",
                              ifelse(Participant %in% prob_responders,"Responder","Non-responder")))

params_base <- full_join(metab_data_scale_base,diversity_all_base) %>% 
  full_join(.,olink_data_base) %>%
  ungroup %>% 
  select(-Timepoint_label)
  #full_join(.,food_data_norm_base)

#16s
merged_tp_chunk_seq_table <- readRDS("~/R/Project_Probiotic/CLEAN_CODE/rds_obj/seq_table_merged_tp_chunk.rds")
ps_tree <- readRDS(file="~/user_data/ProbioticStudy_16S/saved_rds/phyloseq_obj_tree_ProbioticStudy_ppt.rds")
ps_taxa <- ps_tree@tax_table %>% data.frame()
```

LOOCV RFE
```{r}
loocv_rf_function <- function(df) {
  set.seed(200)
  folds <- groupKFold(df$Participant,k=length(unique(df$Participant)))
  control <- rfeControl(functions=rfFuncs, method="LOOCV",number = 100,index=folds)
  df_temp <- select(df,-Participant) %>% as.data.frame
  df_centered <- center_scale(df_temp[,2:dim(df_temp)[2]])
  rfe_obj <- rfe(df_centered,as.factor(df_temp[,1]), sizes=(2:ncol(df_temp)-1), rfeControl=control)
  return(rfe_obj)
}
center_scale <- function(x) {
    scale(x, scale = FALSE)
}
loocv_results_dfs <- function(loocv_object,data_type_char){
  accuracy_df <- data.frame(data_type=data_type_char,
                            max_acc=max(loocv_object[["results"]][["Accuracy"]]),
                            optsize=loocv_object[["optsize"]])
  optVariables_df <- data.frame(data_type=data_type_char,
                            rank_order=c(1:length(loocv_object[["optVariables"]])),
                            optVariables=loocv_object[["optVariables"]])
  return(list(accuracy_df,optVariables_df))
}
```

LOOCV for blood params at baseline 
```{r}
for(i in 3:ncol(params_base)){
  params_base[is.na(params_base[,i]), i] <- mean(params_base[,i], na.rm = TRUE)
}

df_in <- params_base %>% 
  filter(prob_response %in% c("Responder","Non-responder")) %>% 
  select(-treatment_type)

loocv_rf_params_base <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_params_base <- loocv_rf_params_base %>% 
  loocv_results_dfs(.,"params_base") #72% accuracy

loocv_rf_results_params_base_optVar <- loocv_rf_results_params_base[[2]]

params_base_loocv <- params_base %>% 
  ungroup %>% 
  gather(key=param,value=value,-Participant,-treatment_type,-prob_response ) %>% 
  filter(param %in% loocv_rf_results_params_base_optVar[c(1:10),]$optVariables,
         prob_response %in% c("Responder","Non-responder"))

ggplot(params_base_loocv,aes(x=prob_response,y=value,fill=prob_response))+
  geom_boxplot()+
  facet_wrap(~param,scales="free")+
  theme_bw()+
  scale_fill_manual(values=c("#b2182b","#2166ac"))

#food acrokss whole study
df_in <- food_data_study %>% 
  filter(prob_response %in% c("Responder","Non-responder")) %>% 
  ungroup %>% 
  select(-treatment_type,-Timepoint_label)

loocv_rf_food_study <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_food_study <- loocv_rf_food_study %>% 
  loocv_results_dfs(.,"food_study") #60% accuracy

#olink at baseline
df_in <- olink_data_tp_chunk %>% 
  filter(prob_response %in% c("Responder","Non-responder")) %>% 
  ungroup %>% 
  select(Participant,prob_response,ADA:X4E.BP1)

loocv_rf_olink_base <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_olink_base <- loocv_rf_olink_base %>% 
  loocv_results_dfs(.,"olink_base") #60% accuracy

#metabolites at baseline
df_in <- metab_data_scale %>% 
  filter(prob_response %in% c("Responder","Non-responder") & Timepoint==1) %>% 
  ungroup %>% 
  select(-Timepoint,-treatment_type)

loocv_rf_metab_base <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_metab_base <- loocv_rf_metab_base %>% 
  loocv_results_dfs(.,"metab_base") #60% accuracy

#alpha diversity at baseline
df_in <- diversity_all_base %>% 
  filter(prob_response %in% c("Responder","Non-responder")) %>% 
  ungroup %>% 
  select(-Timepoint_label,-treatment_type)

loocv_rf_adiv_base <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_adiv_base <- loocv_rf_adiv_base %>% 
  loocv_results_dfs(.,"adiv_base") #60% accuracy

```

Test metabolic syn params at baseline between probiotic and placebo group
```{r}
#treat
df_in <- params_base %>% 
  ungroup %>% 
  filter(treatment_type %in% c("Probiotic","Placebo")) %>% 
  select(-prob_response)

loocv_rf_params_base_tt <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_params_base_tt <- loocv_rf_params_base_tt %>% 
  loocv_results_dfs(.,"params_base_tt") #68% accuracy

loocv_rf_results_params_base_tt_optVar <- loocv_rf_results_params_base_tt[[2]]

params_base_tt_loocv <- params_base %>%
  ungroup %>%
  gather(key=param,value=value,-Participant,-treatment_type,-prob_response) %>%
  filter(param %in% loocv_rf_results_params_base_tt_optVar[c(1:10),]$optVariables & treatment_type%in%c("Probiotic","Placebo"))

ggplot(params_base_tt_loocv,aes(x=treatment_type,y=value,fill=treatment_type))+
  geom_boxplot()+
  facet_wrap(~param,scales="free")
```

```{r}
# for(i in 3:ncol(food_data_study)){
#   food_data_study[is.na(food_data_study[,i]), i] <- mean(food_data_study[,i], na.rm = TRUE)
# }

df_in <- food_data_study %>% 
  ungroup %>% 
  filter(prob_response %in% c("Responder","Non-responder") & Timepoint_label=="Study") %>% 
  select(-treatment_type,-Timepoint_label)

loocv_rf_food_study <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_food_study <- loocv_rf_food_study %>% 
  loocv_results_dfs(.,"food_study") #63% accuracy

loocv_rf_results_params_base_optVar <- loocv_rf_results_params_base[[2]]

params_base_loocv <- params_base %>% 
  ungroup %>% 
  gather(key=param,value=value,-Participant,-treatment_type,-prob_response,-Timepoint_label) %>% 
  filter(param %in% loocv_rf_results_params_base_optVar$optVariables)

ggplot(params_base_loocv,aes(x=prob_response,y=value,fill=prob_response))+
  geom_boxplot()+
  facet_wrap(~param,scales="free")
```


LOOCV for baseline merged ASVs
```{r}
df_in <- merged_tp_chunk_seq_table %>% 
  filter(prob_response %in% c("Responder","Non-responder") & tp_chunk=="Baseline") %>% 
  select(-treatment_type,-tp_chunk)

loocv_rf_seq_table_base <- df_in %>% 
  loocv_rf_function()
loocv_rf_results_seq_table_base <- loocv_rf_seq_table_base %>% 
  loocv_results_dfs(.,"seq_table") #65% accuracy
#aveRDS(loocv_rf_results_seq_table_base,"~/R/Project_Probiotic/CLEAN_CODE/rds_obj/loocv_rf_seq_table.rds")

loocv_rf_results_seq_table_base_optVar <- loocv_rf_results_seq_table_base[[2]]



```

```{r}
unpairedYSAM <- function(df1,df2){
  y = c(rep(0,dim(df1)[1]),rep(1,dim(df2)[1]))
  return(y)
}

unpairedXSAM <- function(df1,df2,startIDs,endIDs){
  x1 = t(df1[-c(startIDs:endIDs)])
  #print(head(x1))
  x2 = t(df2[-c(startIDs:endIDs)])
  #print(head(x2))
  x_all <- cbind(x1,x2) 
  #print(x_all)
  return(x_all)
}

var_function <- function(df_spread,num_quantile){
  vars_vector <- df_spread %>% 
    as.matrix() %>% 
    colVars()
  name_vector <- df_spread %>% 
    colnames()
  filt_names <- name_vector[vars_vector>quantile(vars_vector,na.rm=TRUE)[num_quantile]] #top 75% varying proteins =2
  return(filt_names)
}
```

Unpaired ttest between responders and non responders during study
  - parameters
  - alpha diversity
  - olink
```{r}
params_base_R <- params_base %>% 
  filter(prob_response=="Responder")
params_base_NR <- params_base %>% 
  filter(prob_response=="Non-responder")
params_base_placebo <- params_base %>% 
  filter(prob_response=="Placebo")

#prob R vs. NR
set.seed(20)
siggenes_input_data <- unpairedXSAM(params_base_R,params_base_NR,1,3)
siggenes_input_cl <- unpairedYSAM(params_base_R,params_base_NR)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
# siggenes_df <- summary(siggenes_output,1.278269)
# siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig_hw360 <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)

olink_base_R <- olink_data_tp_chunk %>% 
  filter(prob_response=="Responder" & Timepoint_label=="Baseline") %>% 
  select(-prob_response)
olink_base_NR <- olink_data_tp_chunk %>% 
  filter(prob_response=="Non-responder" & Timepoint_label=="Baseline") %>% 
  select(-prob_response)
olink_base_placebo <- olink_data_tp_chunk %>% 
  filter(prob_response=="Placebo" & Timepoint_label=="Baseline") %>% 
  select(-prob_response)

#prob R vs. NR
set.seed(20)
siggenes_input_data <- unpairedXSAM(olink_base_R,olink_base_NR,1,3)
siggenes_input_cl <- unpairedYSAM(olink_base_R,olink_base_NR)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
```

```{r}
seq_table_base <- merged_tp_chunk_seq_table %>% 
  filter(tp_chunk=="Baseline" &prob_response!="Placebo")
seq_table_base_raw <- seq_table_base %>% 
  select(-c(Participant:tp_chunk))
dist_seq_table_base <- dist(seq_table_base_raw, method = "euclidean")
ps.adonis <- adonis(dist_seq_table_base ~ prob_response, data = seq_table_base, perm=9999)
as.data.frame(ps.adonis$aov.tab)["prob_response", "Pr(>F)"]

betadisper_prob_resp <- betadisper(dist_seq_table_base, seq_table_base$prob_response)
betadisper_prob_resp_df <- data.frame(Participant=seq_table_base$Participant,
                            treatment_type=seq_table_base$treatment_type,
                            prob_response=seq_table_base$prob_response,
                            dist_cent=betadisper_prob_resp[["distances"]],
                            betadisper_prob_resp[["vectors"]]) 

ggplot(betadisper_prob_resp_df,aes(x=PCoA1,y=PCoA2,colour=prob_response))+
  geom_point(size=3)+
  stat_ellipse(aes(x=PCoA1,y=PCoA2,colour=prob_response),level = 0.50)+
  scale_color_manual(values=c("#b2182b","gray85","#2166ac"))+
  theme_classic()+
  xlab("ASV Profile PCoA1 (45%)")+
  ylab("ASV Profile PCoA2 (26%)")+
  theme(text = element_text(size=18))

physeq_otu <- otu_table(seq_table_base_raw,taxa_are_rows = F)
physeq_samp <- seq_table_base %>% 
  select(Participant:tp_chunk) %>% 
  sample_data()
ps_taxa_base <- ps_tree@tax_table
physeq_base <- phyloseq(physeq_otu,physeq_samp,ps_taxa_base)

ord_obj_base <- ordinate(physeq_base, method = "PCoA", distance = "bray")
ord_obj_base_vectors <- seq_table_base %>% 
  select(Participant:tp_chunk) %>% 
  cbind(.,ord_obj_base$vectors)

dist.uf_base <- phyloseq::distance(physeq_base, method = "bray")
ps.adonis_base <- adonis(dist.uf_base ~ prob_response, data = seq_table_base, perm=9999)
as.data.frame(ps.adonis_base$aov.tab)["prob_response", "Pr(>F)"]

ggplot(ord_obj_base_vectors,aes(x=Axis.1,y=Axis.2,colour=prob_response))+
  geom_point(size=3)+
  stat_ellipse(aes(x=Axis.1,y=Axis.2,colour=prob_response),level = 0.50)+
  scale_color_manual(values=c("#b2182b","#2166ac"))+
  theme_classic()+
  xlab("ASV Profile PCoA1 (10.6%)")+
  ylab("ASV Profile PCoA2 (8%)")+
  theme(text = element_text(size=18))

otu_raw_filt_index <- apply(seq_table_base_raw,2,function(X){sum(X>0)>0.25*length(X)}) %>% #filter asvs to at least 10% of samples
  as.data.frame %>% 
  filter(.==TRUE) %>% 
  rownames

plot_ord_obj_base <- plot_ordination(physeq_base, ord_obj_base,type = "taxa")
loadings_df <- data.frame(plot_ord_obj_base[["plot_env"]][["DF"]]) %>% 
  mutate(ASV=rownames(.),
         prev_level=ifelse(ASV %in% otu_raw_filt_index,"high","low")) %>% 
  filter(prev_level=="high")

loadings_df_grouped <- loadings_df %>% 
  na.omit %>% 
  group_by(Phylum,Class,Order,Family) %>% 
  dplyr::summarize(Axis_1_avg=mean(Axis.1),
                   Axis_2_avg=mean(Axis.2))

ggplot(loadings_df_grouped,aes(x=Axis_1_avg,y=Axis_2_avg)) +
  geom_point()+
  geom_text(aes(label=Family))
```



```{r}


seq_table_base_raw <- merged_tp_chunk_seq_table %>% 
  filter(tp_chunk=="Baseline") %>% 
  select(-c(Participant:tp_chunk))

#filter foods to those non-zero in at least x% of samples
seq_table_base_names <- apply(seq_table_base_raw,2,function(X){sum(X>0,na.rm = T)>0.25*length(X)}) %>% #filter parameters to at least 50% of samples, goes from 109 parameters to 95
  as.data.frame %>% 
  filter(.==TRUE) %>% 
  rownames

seq_table_base_filt <- merged_tp_chunk_seq_table %>% 
  filter(tp_chunk=="Baseline") %>% 
  select(Participant:tp_chunk,!!!seq_table_base_names)

seq_table_base_filt_R <- seq_table_base_filt %>% 
  filter(prob_response=="Responder")
seq_table_base_filt_NR <- seq_table_base_filt %>% 
  filter(prob_response=="Non-responder")
seq_table_base_filt_placebo <- seq_table_base_filt %>% 
  filter(prob_response=="Placebo")

#prob R vs. NR
set.seed(20)
siggenes_input_data <- unpairedXSAM(seq_table_base_filt_R,seq_table_base_filt_NR,1,4)
siggenes_input_cl <- unpairedYSAM(seq_table_base_filt_R,seq_table_base_filt_NR)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.001)
# siggenes_df <- summary(siggenes_output,1.278269)
siggenes_df <- summary(siggenes_output,delta_siggenes[2])
siggenes_mat.sig_seq_table_base <- siggenes_df@mat.sig %>%
  filter(q.value <= 0.1) %>%
  mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
  select(Analyte,q.value)
```


```{r}
food_norm_study_chunk_raw <- food_data_norm_tp_chunk %>% 
  ungroup %>% 
  spread(key=food,value=food_value_norm_med) %>% 
  select(-c(Participant:tp_chunk))

#filter foods to those non-zero in at least x% of samples
hw360_food_names <- apply(food_norm_study_chunk_raw,2,function(X){sum(X>0,na.rm = T)>0.25*length(X)}) %>% #filter parameters to at least 50% of samples, goes from 109 parameters to 95
  as.data.frame %>% 
  filter(.==TRUE) %>% 
  rownames

food_data_study_chunk_resp_int <- food_data_norm_tp_chunk %>% 
  filter(prob_response=="Responder" & tp_chunk=="Study" & (food %in% hw360_food_names)) %>% 
  spread(key=food,value=food_value_norm_med) %>% 
  ungroup %>% 
  na.omit
food_data_study_chunk_NONresp_int <- food_data_norm_tp_chunk %>% 
  filter(prob_response=="Non-responder" & tp_chunk=="Study" & (food %in% hw360_food_names)) %>% 
  spread(key=food,value=food_value_norm_med) %>% 
  ungroup %>% 
  na.omit
food_data_study_chunk_placebo_int <- food_data_norm_tp_chunk %>% 
  filter(prob_response=="Placebo" & tp_chunk=="Study" & (food %in% hw360_food_names)) %>% 
  spread(key=food,value=food_value_norm_med) %>% 
  ungroup %>% 
  na.omit

#prob R vs. NR
set.seed(20)
siggenes_input_data <- unpairedXSAM(food_data_study_chunk_resp_int,food_data_study_chunk_NONresp_int,1,4)
siggenes_input_cl <- unpairedYSAM(food_data_study_chunk_resp_int,food_data_study_chunk_NONresp_int)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
# siggenes_df <- summary(siggenes_output,1.278269)
siggenes_df <- summary(siggenes_output,delta_siggenes[2])
siggenes_mat.sig_hw360 <- siggenes_df@mat.sig %>%
  filter(q.value <= 0.1) %>%
  mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
  select(Analyte,q.value)

#prob vs plac
set.seed(20)
siggenes_input_data <- unpairedXSAM(food_data_study_chunk_resp_int,food_data_study_chunk_placebo_int,1,4)
siggenes_input_cl <- unpairedYSAM(food_data_study_chunk_resp_int,food_data_study_chunk_placebo_int)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
siggenes_df <- summary(siggenes_output,1.149646)
siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig_hw360 <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)

#prob R vs plac
set.seed(20)
siggenes_input_data <- unpairedXSAM(food_data_study_chunk_resp_int,food_data_study_chunk_placebo_int,1,4)
siggenes_input_cl <- unpairedYSAM(food_data_study_chunk_resp_int,food_data_study_chunk_placebo_int)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
siggenes_df <- summary(siggenes_output,1.166037)
siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig_hw360 <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)

#prob NR vs plac
set.seed(20)
siggenes_input_data <- unpairedXSAM(food_data_study_chunk_NONresp_int,food_data_study_chunk_placebo_int,1,4)
siggenes_input_cl <- unpairedYSAM(food_data_study_chunk_NONresp_int,food_data_study_chunk_placebo_int)
siggenes_output <- siggenes::sam(data=siggenes_input_data,cl=siggenes_input_cl)
summary(siggenes_output)
delta_siggenes <- findDelta(siggenes_output, fdr = 0.05)
# siggenes_df <- summary(siggenes_output,1.149646)
# siggenes_df <- summary(siggenes_output,delta_siggenes[2])
# siggenes_mat.sig_hw360 <- siggenes_df@mat.sig %>%
#   filter(q.value <= 0.1) %>%
#   mutate(Analyte = rownames(siggenes_df@mat.sig)[c(1:dim(.)[1])]) %>%
#   select(Analyte,q.value)

#plot differences in prob R vs. NR
food_data_study_chunk_sigg_RvNR <- food_data_norm_tp_chunk %>% 
  filter(tp_chunk=="Study" & (food %in% siggenes_mat.sig_hw360$Analyte)) %>% 
  mutate(prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE)) %>% 
  spread(key=food,value=food_value_norm_med) %>%
  na.omit# %>% 
  # mutate_if(is.numeric, scale) 
food_data_study_chunk_sigg_RvNR_mean_summary <- food_data_study_chunk_sigg_RvNR %>% 
  gather(key=sigg_food,value=food_value,-Participant,-treatment_type,-prob_response,-tp_chunk) %>% 
  group_by(prob_response,tp_chunk,sigg_food) %>% 
  dplyr::summarize(prob_resp_avg=mean(food_value,na.rm=TRUE)) %>% 
  spread(key=prob_response,value=prob_resp_avg)
#write.csv(food_data_study_chunk_sigg_RvNR_mean_summary,"~/R/Project_Probiotic/CLEAN_CODE/data/siggenes_food_avg_all_grps.csv")


ind <- sapply(food_data_study_chunk_sigg_RvNR, is.numeric)
food_data_study_chunk_sigg_RvNR[ind] <- lapply(food_data_study_chunk_sigg_RvNR[ind], scale)

food_data_study_chunk_sigg_RvNR_gather <- food_data_study_chunk_sigg_RvNR %>% 
  gather(key=food,value=food_value_norm_med,-c(Participant:tp_chunk))
ggplot(food_data_study_chunk_sigg_RvNR_gather,aes(x=food,y=food_value_norm_med,fill=prob_response))+
  geom_boxplot()+
  # facet_wrap(~food,scales="free")+
  theme_bw()+
  scale_fill_manual(values=c("gray","#b2182b","#2166ac"))+
  theme(text = element_text(size=18),axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Scaled Nutrition Intake")+
  xlab("Treatment Arm/Response Group")
#ggsave(paste(save_figure_path,"food_siggenes_probRvNR_boxplots.pdf"),width=18.2,height = 9.35)

#plot a subset of the sig foods that are less redundant 
food_sig_short <- c("Total.Sugars..g.","Added.Sugars..by.Total.Sugars...g.","Lactose..g.","Total.Folate..mcg.","Calcium..mg.","Sodium..mg.","Zinc..mg.")

food_data_study_chunk_sigg_RvNR_gather_short <- food_data_study_chunk_sigg_RvNR_gather %>% 
  filter(food%in%food_sig_short) %>% 
  mutate(food=factor(food,levels=c("Total.Sugars..g.","Added.Sugars..by.Total.Sugars...g.","Lactose..g.","Total.Folate..mcg.","Calcium..mg.","Sodium..mg.","Zinc..mg."))) %>% 
  arrange(food)


ggplot(food_data_study_chunk_sigg_RvNR_gather_short,aes(x=food,y=food_value_norm_med,fill=prob_response))+
  geom_boxplot()+
  # facet_wrap(~food,scales="free")+
  theme_bw()+
  scale_fill_manual(values=c("gray","#b2182b","#2166ac"))+
  theme(text = element_text(size=20),axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Scaled Nutrition Intake")+
  xlab("Treatment Arm/Response Group")
#ggsave(paste(save_figure_path,"food_siggenes_probRvNR_boxplots_short.pdf"),width=18.2,height = 9.35)

```

Plot non scaled values for supp
```{r}
food_data_study_chunk_sigg_RvNR_noscale_noscale <- food_data_norm_tp_chunk %>% 
  filter(tp_chunk=="Study" & (food %in% siggenes_mat.sig_hw360$Analyte)) %>% 
  mutate(prob_response=factor(prob_response,levels=c("Placebo","Non-responder","Responder"),ordered=TRUE)) %>% 
  spread(key=food,value=food_value_norm_med) %>%
  na.omit# %>% 

food_data_study_chunk_sigg_RvNR_noscale_gather <- food_data_study_chunk_sigg_RvNR_noscale %>% 
  gather(key=food,value=food_value_norm_med,-c(Participant:tp_chunk))
ggplot(food_data_study_chunk_sigg_RvNR_noscale_gather,aes(x=food,y=food_value_norm_med,fill=prob_response))+
  geom_boxplot()+
  # facet_wrap(~food,scales="free")+
  theme_bw()+
  scale_fill_manual(values=c("gray","#b2182b","#2166ac"))+
  theme(text = element_text(size=18),axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Scaled Nutrition Intake")+
  xlab("Treatment Arm/Response Group")
#ggsave(paste(save_figure_path,"food_siggenes_probRvNR_boxplots.pdf"),width=18.2,height = 9.35)

#plot a subset of the sig foods that are less redundant 
food_sig_short <- c("Total.Sugars..g.","Added.Sugars..by.Total.Sugars...g.","Lactose..g.","Total.Folate..mcg.","Calcium..mg.","Sodium..mg.","Zinc..mg.")

food_data_study_chunk_sigg_RvNR_noscale_gather_short <- food_data_study_chunk_sigg_RvNR_noscale_gather %>% 
  filter(food%in%food_sig_short) %>% 
  mutate(food=factor(food,levels=c("Total.Sugars..g.","Added.Sugars..by.Total.Sugars...g.","Lactose..g.","Total.Folate..mcg.","Calcium..mg.","Sodium..mg.","Zinc..mg."))) %>% 
  arrange(food)

ggplot(food_data_study_chunk_sigg_RvNR_noscale_gather_short,aes(x=food,y=food_value_norm_med,fill=prob_response))+
  geom_boxplot()+
  facet_wrap(~food,scales="free")+
  theme_bw()+
  scale_fill_manual(values=c("gray","#b2182b","#2166ac"))+
  theme(text = element_text(size=20))+
  ylab("Scaled Nutrition Intake")+
  xlab("Treatment Arm/Response Group")
#ggsave(paste(save_figure_path,"food_siggenes_probRvNR_boxplots_short.pdf"),width=18.2,height = 9.35)
```

